---
title: "fig1 code"
output: html_document
---
#input packages
```{r setup, include=FALSE}
library(GGally)
library("FactoMineR")
library("factoextra")
#install.packages("pals")
library(pals)
library(ggplot2)
library(ggfortify)
library(corrplot)
library(data.table)
```
#1data
```{r}
#1get the communtiy level trait
us_eu_CWM=read.csv(".../US.EU.CWM_trait.txt",sep="\t")
names(us_eu_CWM)

#get the specie level trait
us_eu_sp=read.csv(".../US.EU.species_level_trait.txt",sep="\t")
names(us_eu_sp)
names(us_eu_CWM)[4:11]<-c("-P50","Slope","-TLP","LS","Ks","WD","LMA","N")
names(us_eu_sp)[4:11]<-c("-P50","Slope","-TLP","LS","Ks","WD","LMA","N")
#only include the community which the dominant group has higher than 50% 
us_eu_CWM<-us_eu_CWM[us_eu_CWM$per>=0.5,]
#dplyr::count(us_eu_CWM,group)
```
##PCA at communtiy level

```{r fig.width=8,fig.height=5.5}
# PCA with prcomp
m1 <- prcomp(us_eu_CWM[,c(4:11)], center = TRUE,scale. = T,retx=TRUE)
m1$x[,1] <- - m1$x[,1]
m1$rotation[,1] <- -m1$rotation[,1]
autoplot(m1,loadings.label=TRUE,loadings=TRUE)

# Get principal loadings of first two components
us_eu_CWM$PC1 <- predict(m1, newdata = us_eu_CWM)[,1]
us_eu_CWM$PC2 <- predict(m1, newdata = us_eu_CWM)[,2]
# Get arrow end point locations (loadings*2 for effect)
l.x <- m1$rotation[,1]*9
l.y <- m1$rotation[,2]*9
# Get label positions (%15 further than end of arrows)
l.posx <- l.x*1.25
l.posy <- l.y*1.1
# Get labels for plot (variable names)
l.labels <- row.names(m1$rotation)
us_eu_CWM$group<-as.factor(us_eu_CWM$group)
levels(us_eu_CWM$group)
```


#2.species level


```{r }
sp_pca <- prcomp(us_eu_sp[,c(4:11)],center = TRUE,scale. = T,retx=TRUE) 

# Get principal loadings of first two components
us_eu_sp$PC1 <- predict(sp_pca, newdata = us_eu_sp)[,1]
us_eu_sp$PC2 <- predict(sp_pca, newdata = us_eu_sp)[,2]
# Get arrow end point locations (loadings*2 for effect)
l.x <- sp_pca$rotation[,1]*10
l.y <- sp_pca$rotation[,2]*11
# Get label positions (%15 further than end of arrows)
l.posx <- l.x*1.25
l.posy <- l.y*1.1
# Get labels for plot (variable names)
l.labels <- row.names(sp_pca$rotation)
us_eu_sp$group<-as.factor(us_eu_sp$group)
```

#3 fig s3 contribution
##3.1 community-level
```{r}
fviz_eig(cwm_pca, geom="bar",width=1,addlabels = TRUE)
fig.s3a<-fviz_eig(cwm_pca, width=0.2,addlabels = F, ylim = c(0, 50), geom = c("bar","line"),
  barfill = "gray86",
  barcolor = "black",
  linecolor = "red", 
  alpha=0.8,
  hjust = -0.2,
  xlab="PCs",ylab="Explained variances",mian=NULL,
  ggtheme = theme_classic()+
  theme(panel.grid.major.x = element_blank() , 
        axis.ticks.length=unit(0.08,"inch"),
        title = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_text(colour = "black",  size = 12),
        axis.text.x = element_text(colour = "black",  size = 12),
        axis.title.x = element_blank()))
  
fig.s3a
#ggsave(fig2c,filename = '.../fig s3a.tiff',compression='lzw', width = 6,height = 5,dpi=200)
```

##3.2 species-level
```{r}
fviz_eig(sp_pca, geom="bar",width=1,addlabels = TRUE)
fig.s3b<-fviz_eig(sp_pca, width=0.2,addlabels = F, ylim = c(0, 50), geom = c("bar","line"),
  barfill = "gray86",
  barcolor = "black",
  linecolor = "red", 
  alpha=0.8,
  hjust = -0.2,
  xlab="PCs",ylab="Explained variances",mian=NULL,
  ggtheme = theme_classic()+
  theme(panel.grid.major.x = element_blank() , 
        axis.ticks.length=unit(0.08,"inch"),
        title = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_text(colour = "black",  size = 12),
        axis.text.x = element_text(colour = "black",  size = 12),
        axis.title.x = element_blank()))
  
fig.s3b
ggsave(fig.s3c,filename = '.../fig.s3b.tiff',compression='lzw', width = 6,height = 5,dpi=200)
```
#4. fig s4 for PFT
##4.1 community-level
###PC1
```{r}
PC1_com<-ggplot(data = us_eu_CWM, aes(x = group, y = PC1,fill=group)) +
    geom_violin(trim=FALSE,scale = "width")+
    scale_y_continuous(name=expression('PC1 scores'),
          limits=c(-8,9),breaks=seq(-8,8,8))+
    geom_boxplot(width=0.15, fill="white")+
    scale_fill_manual(values=c( "#2166AC","forestgreen","#D6604D"),label=c('BD','BE','NC'))+
    annotate("text", x = 1, y = 8.6, label = "a",size=6)+
    annotate("text", x = 2, y = 7.95, label = "b",size=6)+
    annotate("text", x = 3, y = 5.8, label = "c",size=6)+
    theme_classic()+
    theme(plot.background=element_blank(),
          axis.ticks.length=unit(0.08,"inch"),
          panel.grid.major.y = element_line(size = 0.6 , colour = "grey90") ,
          axis.line=element_line(colour='black',linetype=1),
          legend.position = "none",
          axis.title.y = element_text(colour='black',size=14,vjust=1),
          axis.text.y  = element_text(colour='black',vjust=0.5, size=12),
          axis.title.x = element_blank(),
          axis.text.x  = element_text(colour='black',angle=42, vjust=0.55, size=14))

PC1_com
#ggsave(PC1_com,filename = '.../fig3a.tiff',compression='lzw', width = 4.5,height = 3,dpi=200)
```
###PC2
```{r}
data.lm <- lm(PC2 ~ group, 
              data = us_eu_CWM)
data.av <- aov(data.lm)
summary(data.av)
data.test <- TukeyHSD(data.av)
data.test

PC2_com<-ggplot(data = us_eu_CWM, aes(x = group, y = PC2,fill=group)) +
    geom_violin(trim=FALSE,scale = "width")+
    scale_y_continuous(name=expression('PC2 scores'),
          limits=c(-8,9),breaks=seq(-8,8,8))+
    geom_boxplot(width=0.15, fill="white")+
    scale_fill_manual(values=c( "#2166AC","forestgreen","#D6604D"),label=c('BD','BE','NC'))+
    annotate("text", x = 1, y = 7, label = "a",size=6)+
    annotate("text", x = 2, y = 8, label = "b",size=6)+
    annotate("text", x = 3, y = 5.5, label = "c",size=6)+
    theme_classic()+
    theme(plot.background=element_blank(),
          axis.ticks.length=unit(0.08,"inch"),
          panel.grid.major.y = element_line(size = 0.6 , colour = "grey90") ,
          axis.line=element_line(colour='black',linetype=1),
          legend.position = "none",
          axis.title.y = element_text(colour='black',size=14,vjust=1),
          axis.text.y  = element_text(colour='black',vjust=0.5, size=12),
          axis.title.x = element_blank(),
          axis.text.x  = element_text(colour='black',angle=42, vjust=0.55, size=14))

PC2_com
#ggsave(PC2_com,filename = '.../PC2_com.tiff',compression='lzw', width = 4.5,height = 3,dpi=200)
```
##4.2 species-level

###pc1
```{r}
data.lm <- lm(PC1 ~ group, 
              data = us_eu_sp)
data.av <- aov(data.lm)
summary(data.av)
data.test <- TukeyHSD(data.av)
data.test

PC1_sp<-ggplot(data = us_eu_sp, aes(x = group, y = PC1,fill=group)) +
    geom_violin(trim=FALSE,scale = "width")+
    scale_y_continuous(name=expression('PC1 scores'),
          limits=c(-8,9),breaks=seq(-8,8,8))+
    geom_boxplot(width=0.15, fill="white")+
    scale_fill_manual(values=c( "#2166AC","forestgreen","#D6604D"),label=c('BD','BE','NC'))+
    annotate("text", x = 1, y = 7, label = "a",size=6)+
    annotate("text", x = 2, y = 7, label = "a",size=6)+
    annotate("text", x = 3, y = 5, label = "b",size=6)+
    theme_classic()+
    theme(plot.background=element_blank(),
          axis.ticks.length=unit(0.08,"inch"),
          panel.grid.major.y = element_line(size = 0.6 , colour = "grey90") ,
          axis.line=element_line(colour='black',linetype=1),
          legend.position = "none",
          axis.title.y = element_text(colour='black',size=14,vjust=1),
          axis.text.y  = element_text(colour='black',vjust=0.5, size=12),
          axis.title.x = element_blank(),
          axis.text.x  = element_text(colour='black',angle=42, vjust=0.55, size=14))

PC1_sp
#ggsave(PC1_sp,filename = '.../PC1_sp.tiff',compression='lzw', width = 4.5,height = 3,dpi=200)
```
###PC2
```{r}
data.lm <- lm(PC2 ~ group, 
              data = us_eu_sp)
data.av <- aov(data.lm)
summary(data.av)
data.test <- TukeyHSD(data.av)
data.test

PC2_sp<-ggplot(data = us_eu_sp, aes(x = group, y = PC2,fill=group)) +
    geom_violin(trim=FALSE,scale = "width")+
    scale_y_continuous(name=expression('PC2 scores'),
          limits=c(-8,9),breaks=seq(-8,8,8))+
    geom_boxplot(width=0.15, fill="white")+
    scale_fill_manual(values=c("#2166AC","forestgreen", "#D6604D"),label=c('BD','BE','NC'))+
    annotate("text", x = 1, y = 6, label = "a",size=6)+
    annotate("text", x = 2, y = 7, label = "b",size=6)+
    annotate("text", x = 3, y = 5, label = "c",size=6)+
    theme_classic()+
    theme(plot.background=element_blank(),
          axis.ticks.length=unit(0.08,"inch"),
          panel.grid.major.y = element_line(size = 0.6 , colour = "grey90") ,
          axis.line=element_line(colour='black',linetype=1),
          legend.position = "none",
          axis.title.y = element_text(colour='black',size=14,vjust=1),
          axis.text.y  = element_text(colour='black',vjust=0.5, size=12),
          axis.title.x = element_blank(),
          axis.text.x  = element_text(colour='black',angle=42, vjust=0.55, size=14))

PC2_sp
#ggsave(PC2_sp,filename = '.../PC2_sp.tiff',compression='lzw', width = 4.5,height = 3,dpi=200)
```




#5 test PCA excluding lots <2
```{r}
#get the scommunity which have plots more than 2 plots
us_eu_CWM_test<-us_eu_CWM[us_eu_CWM$plot.n>2,]

```

```{r}
cwm_pca <- prcomp(us_eu_CWM_test[,c(4:11)],center = TRUE,scale. = T,retx=TRUE) 


cwm_pca$x[,1] <- - cwm_pca$x[,1]
cwm_pca$rotation[,1] <- -cwm_pca$rotation[,1]
autoplot(cwm_pca,loadings.label=TRUE,loadings=TRUE)


fviz_pca_biplot(cwm_pca, axes = c(1, 2),
                    pointsize = 1,alpha=0.5,
                    col.ind = us_eu_CWM_test$group, 
                    palette = c("#2166AC","forestgreen", "#D6604D"),
                    addEllipses = F, label = "var",
                    col.var = "black", repel = TRUE,arrowsize = 1.2, alpha_arrow = 1,
                    ggtheme = theme_minimal(),
                    legend.title = "PFTs")
```
##fig.s6a
```{r fig.width=8,fig.height=5.5}
# PCA with prcomp
m1 <- prcomp(us_eu_CWM_test[,c(4:11)], center = TRUE,scale. = T,retx=TRUE)
m1$x[,1] <- - m1$x[,1]
m1$rotation[,1] <- -m1$rotation[,1]

m1$x[,2] <- - m1$x[,2]
m1$rotation[,2] <- -m1$rotation[,2]
autoplot(m1,loadings.label=TRUE,loadings=TRUE)

# Get principal loadings of first two components
us_eu_CWM_test$PC1 <- predict(m1, newdata = us_eu_CWM_test)[,1]
us_eu_CWM_test$PC2 <- predict(m1, newdata = us_eu_CWM_test)[,2]
# Get arrow end point locations (loadings*2 for effect)
l.x <- m1$rotation[,1]*9
l.y <- m1$rotation[,2]*9
# Get label positions (%15 further than end of arrows)
l.posx <- l.x*1.25
l.posy <- l.y*1.1
# Get labels for plot (variable names)
l.labels <- row.names(m1$rotation)
us_eu_CWM_test$group<-as.factor(us_eu_CWM_test$group)
levels(us_eu_CWM_test$group)
# Plot with grouping by flower
fig.s6a<-ggplot() +
  geom_point(data = us_eu_CWM_test, aes(PC1, PC2, color=group),alpha=0.6,shape=15) +
  geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.5)+
  geom_vline(xintercept=0, linetype="dashed", color = "black", size=0.5)+
  scale_color_manual(name  ="Forest type",values = c( "#2166AC","forestgreen","#D6604D")) + 
  scale_size_continuous(name  ="PFT proportion",range = c(0, 2))+
  geom_segment(aes(x=0, y=0, xend = l.x, yend = l.y), 
               arrow = arrow(length = unit(0.4, "cm")),color = "black") +
  #geom_text(aes(x = l.posx, y = l.posy, label = l.labels), color = "black", size = 6, hjust = 0) +
  
  scale_x_continuous(limits=c(-8,8),breaks=seq(-8,8,4)) +
  scale_y_continuous(limits=c(-8,8),breaks=seq(-8,8,4))+
  ylab("PC2 (16.7%)")+xlab("PC1 (45.2%)")+
  theme(
    panel.background = element_rect(color="black",fill="white"), 
    axis.text.y = element_text(colour = "black",  size = 14),
    axis.text.x = element_text(colour = "black",  size = 14),
    axis.title.x = element_text(colour = "black",  size = 16),
    axis.title.y = element_text(colour = "black",  size = 16),
    legend.key=element_blank(),
  legend.text = element_text(colour="black", size = 12),
  legend.title = element_text(size=14),
  legend.key.size = unit(0.7, 'cm'),
  legend.background = element_blank(),
  text = element_text(size = 12),
  legend.position= c(0.15,0.15),
  axis.ticks.length.y = unit(.3, "cm"),
  axis.ticks.length.x = unit(.3, "cm")
  )+guides(color=guide_legend(title="Forest type")) 
fig.s6a
#ggsave(plot=fig.s6a,filename = '.../fig.s6a.tiff', ,compression='lzw',  width = 10,height = 6,dpi=1200,units = "cm")
```
##fig. s6b
```{r}
var<-get_pca_var(cwm_pca)
library("corrplot")
corrplot(var$contrib, is.corr=FALSE)
pca12_var<-data.table(var$contrib)
Trait<-c("-P50","Slope","-TLP","LS","Ks","WD","LMA","N")
pca12_var<-cbind(pca12_var,Trait)

levels(as.factor(pca12_var$Trait))
pca12_var$Trait = factor(pca12_var$Trait, levels=c("WD","-TLP","Slope","-P50",
                                                   "LMA","LS","Ks","N"))
d1<-ggplot(pca12_var,aes(Trait,Dim.1))+geom_bar(stat = "identity",  width=0.8,
                                            alpha=0.8,fill = "gray80",color="black") +
    scale_y_continuous(limits=c(0,30),breaks=seq(0,30,15))+
  coord_flip() +labs(y="PCA 1", x="Functional traits")+theme_classic()+
  
  theme(panel.grid.major.x = element_blank() , 
        axis.ticks.length=unit(0.08,"inch"),axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(colour = "black",  size = 12),
        axis.title.x = element_blank())
d1

d2<-ggplot(pca12_var,aes(Trait,Dim.2))+geom_bar(stat = "identity",  width=0.8,
                                            alpha=0.8,fill = "gray80",color="black") +
    scale_y_continuous(limits=c(0,50),breaks=seq(0,50,25))+
  coord_flip() +labs(y="PCA 2", x="Functional traits")+theme_classic()+
  theme(panel.grid.major.x = element_blank() , 
        axis.ticks.length=unit(0.08,"inch"),axis.title.y = element_blank(),
                axis.text.y = element_blank(),
        axis.text.x = element_text(colour = "black",  size = 12),
        axis.title.x = element_blank())
d2
#make the figure s6b
library(ggpubr)
combined_plot <- ggarrange(d1,d2,nrow = 1,ncol = 2)
combined_plot
ggsave(combined_plot,filename = '.../fig.s6b.tiff',compression='lzw', width = 3.4,height = 4.2,dpi=200)
```

##fig.s6c
```{r}
names(us_eu_CWM_test)
corrplot(corr=cor(us_eu_CWM_test[,c(11,8,7,10,4,5,6,9)]),order="original",type="upper",
         tl.pos ="d",method = "number")
corrplot.mixed(corr=cor(us_eu_CWM_test[,c(11,8,7,10,4,5,6,9)]),order="original",lower="ellipse",
               upper="number",tl.pos =c("d"))
```

```{r fig.width=12,fig.height=4}
cor_metric_cwm<-us_eu_CWM_test[,c("N","Ks","LS","LMA","-P50","Slope","-TLP","WD")]
corrplot.mixed(corr=cor(cor_metric_cwm), order="original",lower="ellipse",
               upper="number",tl.pos =c("d"),insig = "p-value",sig.level = 0.001)
```
